\begin{thebibliography}{10}

\bibitem{rabiner1993}
L.~Rabiner and B.~Juang, \emph{Fundamentals of Speech Recognition}. Englewood Cliffs, NJ, USA: Prentice-Hall, 1993.

\bibitem{davis1980}
S.~Davis and P.~Mermelstein, "Comparison of parametric representations for monosyllabic word recognition in continuously spoken sentences," \emph{IEEE Transactions on Acoustics, Speech, and Signal Processing}, vol. 28, no. 4, pp. 357--366, Aug. 1980.

\bibitem{oppenheim2010}
A.~V. Oppenheim and R.~W. Schafer, \emph{Discrete-Time Signal Processing}, 3rd ed. Upper Saddle River, NJ, USA: Pearson, 2010.

\bibitem{campbell1997}
J.~P. Campbell, "Speaker recognition: a tutorial," \emph{Proceedings of the IEEE}, vol. 85, no. 9, pp. 1437--1462, Sept. 1997.

\bibitem{kinnunen2010}
T.~Kinnunen and H.~Li, "An overview of text-independent speaker recognition: From features to supervectors," \emph{Speech Communication}, vol. 52, no. 1, pp. 12--40, 2010.

\bibitem{itakura1975}
F.~Itakura, "Minimum prediction residual principle applied to speech recognition," \emph{IEEE Transactions on Acoustics, Speech, and Signal Processing}, vol. 23, no. 1, pp. 67--72, Feb. 1975.

\bibitem{reynolds1995}
D.~A. Reynolds, "Speaker identification and verification using Gaussian mixture speaker models," \emph{Speech Communication}, vol. 17, no. 1-2, pp. 91--108, 1995.

\bibitem{bengio2007}
Y.~Bengio, P.~Lamblin, D.~Popovici, and H.~Larochelle, "Greedy layer-wise training of deep networks," \emph{Advances in Neural Information Processing Systems}, vol. 19, 2007.

\bibitem{welch1967}
P.~Welch, "The use of fast Fourier transform for the estimation of power spectra: A method based on time averaging over short, modified periodograms," \emph{IEEE Transactions on Audio and Electroacoustics}, vol. 15, no. 2, pp. 70--73, June 1967.

\bibitem{dehak2011}
N.~Dehak, P.~J. Kenny, R.~Dehak, P.~Dumouchel, and P.~Ouellet, "Front-end factor analysis for speaker verification," \emph{IEEE Transactions on Audio, Speech, and Language Processing}, vol. 19, no. 4, pp. 788--798, May 2011.

\end{thebibliography}